Collocated Snapshot Backup with VM-Centric Deduplication in the Cloud

### 2014-01-22 snapshot_system/fast14_conclusion.tex
added tradeoffs to intro and conclusion, added VO with extra replication approach

\section{Conclusion}
\label{sect:conclusion}
In this paper we propose a collocated backup service built on
the top of a cloud cluster to reduce network traffic and infrastructure.
The key contribution is a VM-centric deduplication scheme to 
maximize fault isolation while delivering competitive deduplication efficiency.
Similarity guided local search reduces cross-VM data dependency and exposes more parallelism  
while global deduplication with a small common data set eliminates popular duplicates.
VM-specific file block packing also enhances fault tolerance by reducing data dependencies.
The design places a special consideration for low-resource usages as a collocated cloud service.
Evaluation using  VM backup data shows that VC strikes a tradeoff and 
can accomplish 96.33\% or 96.9\% of what complete global
deduplication can do.  The availability of snapshots increases substantially with 
a small replication overhead for popular inter-VM chunks.

>>>>

\section{Conclusion}
\label{sect:conclusion}
In this paper we propose a collocated backup service built on
the top of a cloud cluster to reduce network traffic and infrastructure requirements.
The key contribution is a VM-centric deduplication scheme to 
maximize fault isolation with only a slight tradeoff in deduplication efficiency.
Similarity guided local search reduces cross-VM data dependency and exposes more parallelism  
while global deduplication with a small common data set eliminates popular duplicates.
VM-specific file block packing also enhances fault tolerance by reducing data dependencies.
The design places a special consideration for low-resource usages as a collocated cloud service.
Evaluation using  VM backup data shows that VC strikes a tradeoff and 
can accomplish 96.33\% or 96.9\% of what complete global
deduplication can do.  The availability of snapshots increases substantially with 
a small replication overhead for popular inter-VM chunks.

### 2014-01-22 snapshot_system/fast14_deduplication.tex
re
revisions
added tradeoffs to intro and conclusion, added VO with extra replication approach

\subsection{Key VC  Strategies}
\label{sect:vc-strategies}
%\begin{itemize}
%\item 
\textbf{VM-specific local duplicate search within similar segments.}
We start with the changed block tracking approach in a coarse grain segment level.
In our implementation with Xen on an Alibaba platform, the segment size is 2MB
and the device driver is extended to support tracking changed segments using a dirty bitmap. 
Since every write for a segment will touch a dirty bit, the device driver maintains dirty bits in memory
and cannot afford a small segment size.
It should be noted that dirty bit tracking is supported or can be easily implemented in 
major virtualization solution vendors. For example,
the VMWare hypervisor has an API to let external backup applications know 
the changed areas since last backup. 
The Microsoft SDK provides an API that allows external applications to monitor 
the VM's I/O traffic and implement such changed block tracking feature.

@@

% {\it [Need to find a place to put these numbers in: Total number of chunks 
% in 350 snapshots: 1,546,635,485. 
% Total number of chunks after localized dedup: 283,121,924. Total number of unique chunks: 87,692,682.]}
\subsection{Impact on Deduplication Efficiency}
Choosing the value $k$ for the most popular chunks affects the deduplication efficiency.
We analyze this impact based on the characteristics  of the VM snapshot traces
studied from  application datasets.
A previous study shows that the popularity of data chunks after local deduplication follows 
a Zipf-like distribution\cite{Breslau1999a} and its
exponent $\alpha$ is ranged between 0.65  and  0.7~\cite{WeiZhangIEEE}. 
Figure~\ref{fig:Datazipf} illustrates the Zipf-like distribution of chunk popularity.
The parameters we will use in our analysis below are defined in
Table~\ref{tab:symbol}.

\begin{table}[htbp]
\centering
\tabcolsep=0.11cm
\begin{tabular}{|p{1.0cm}|p{5.75cm}|}
\hline
$k$ &  the number of top most popular chunks selected for deduplication\\ 
\hline
$c$ &  the total amount of data chunks in a cluster of VMs\\ 
\hline
$c_u$ &  the total amount of unique fingerprints after perfect  deduplication\\
\hline
$f_i$ &  the frequency for the $i$th most popular fingerprint\\
\hline
$\delta$ &  the percentage of duplicates detected in local deduplication\\
\hline
$\sigma$ & =$\frac{k}{c_u}$ which is  the percentage of unique data  belonging to  PDS\\
\hline
$p$ & the number of machines in the cluster\\
\hline
$V$ & the average number of VMs hosted on each machine\\
\hline
$E_c, E_o$ & deduplication efficiency of VC and VO \\
%\hline
%$D$ & the amount of unique data on each machine\\
%\hline
%$s$ & the average data chunk size. Our setting is  4K.\\
\hline
$s$ & the average number of chunks per FSB\\
%\hline
%$m$ & memory size on each node used by VC\\ 
%\hline
%$E$ & the size of an popular data index entry\\
\hline
$N_1$ & the average number  of non-PDS FSBs blocks in a VM for VC\\
\hline
$N_2$ & the average number  of PDS FSBs in a VM for VC\\
\hline
$N_o$ & the average number  of FSBs  in a VM for VO\\
\hline
A(r) & the availability of an FSB  with replication degree $r$\\
\hline
\end{tabular}
\caption{Modeling  parameters}
\label{tab:symbol}
\end{table}

@@

@@

Now we  assess  the impact of losing $d$ machines 
to the VC and VO approaches.  
A large $r_c/r$ ratio can have a positive impact on full availability of VM snapshot blocks.
We use an FSB rather than a deduplication
data chunk as our unit of failure because the DFS keeps
file system blocks as its base unit of storage.
To compute the full availability of all snapshots of a VM, we derive
the probability of losing a snapshot FSB of a VM by
estimating the number of file system blocks per VM in each approach.
As illustrated in Figure~\ref{fig:share},
we build a bipartite graph representing the association from unique file system blocks
to their corresponding VMs in each approach. An association edge is  drawn  from an FSB to a VM 
if this block is used by the VM.

@@

\item
When $r_c \leq d$, both non-PDS and PDS file system blocks in VC can have a loss.
The full snapshot availability of  a VM in the VC approach is
%\begin{equation}
%\label{eq:VC2}
% (1-\frac{ \binom{d}{r}} { \binom{p}{r} })^{N_1} 
% *
% (1-\frac{ \binom{d}{r_c}} { \binom{p}{r_c} })^{N_2}.
\[
A(r)^{N_1} * A(r_c)^{N_2}
\]
%\end{equation}
%comparing  Formula~\ref{eq:VO} and~\ref{eq:VC}. 
\end{itemize} 
%In the evaluation discussed in Section~\ref{sect:evaldedup},
We have considered a worst case scenario that
every PDS FSB is shared by all VMs in the VC approach, which leads to a large $N_2$ value. 
Even with that, the availability of VC snapshots is still much higher than VO and  
there are two reasons for this:  1) $N_1$ is much smaller than $N_o$ as discussed previously.
2)  $A(r) < A(r_c)$ because $r < r_c$.  
Table~\ref{tab:fsb-availability} lists the $A(r)$ values with
%that the availability of an individual file system block
different replication degrees, to demonstrate the gap between  $A(r)$ and  $A(r_c)$.

>>>>

\subsection{Key VC  Strategies}
\label{sect:vc-strategies}
%\begin{itemize}
%\item 
\textbf{VM-specific local duplicate search within similar segments.}
We start with the changed block tracking approach in a coarse grain segment level.
In our implementation with Xen on an Alibaba platform, the segment size is 2MB
and the device driver is extended to support tracking changed segments using a dirty bitmap. 
Since every write for a segment will touch a dirty bit, the device driver maintains dirty bits in memory
and cannot afford a small segment size.
It should be noted that dirty bit tracking is supported or can be easily implemented in 
major virtualization solution vendors. For example,
the VMWare hypervisor has an API to let external backup applications know 
the changed areas since last backup. 
The Microsoft SDK provides an API that allows external applications to monitor 
the VM's I/O traffic and implement such changed block tracking feature.
%The dirty bit tracking is valuable because it immediately cuts down the number of blocks
%which must be processed, and reduces memory, network, and cpu resource requirements.

@@

% {\it [Need to find a place to put these numbers in: Total number of chunks 
% in 350 snapshots: 1,546,635,485. 
% Total number of chunks after localized dedup: 283,121,924. Total number of unique chunks: 87,692,682.]}

\begin{table}[htbp]
\centering
\tabcolsep=0.11cm
\begin{tabular}{|p{1.0cm}|p{5.75cm}|}
\hline
$k$ &  the number of top most popular chunks selected for deduplication\\ 
\hline
$c$ &  the total amount of data chunks in a cluster of VMs\\ 
\hline
$c_u$ &  the total amount of unique fingerprints after perfect  deduplication\\
\hline
$f_i$ &  the frequency for the $i$th most popular fingerprint\\
\hline
$\delta$ &  the percentage of duplicates detected in local deduplication\\
\hline
$\sigma$ & =$\frac{k}{c_u}$ which is  the percentage of unique data  belonging to  PDS\\
\hline
$p$ & the number of machines in the cluster\\
\hline
$V$ & the average number of VMs per machine\\
\hline
$E_c, E_o$ & deduplication efficiency of VC and VO \\
%\hline
%$D$ & the amount of unique data on each machine\\
%\hline
%$s$ & the average data chunk size. Our setting is  4K.\\
\hline
$s$ & the average number of chunks per FSB\\
%\hline
%$m$ & memory size on each node used by VC\\ 
%\hline
%$E$ & the size of an popular data index entry\\
\hline
$N_1$ & the average number  of non-PDS FSBs blocks in a VM for VC\\
\hline
$N_2$ & the average number  of PDS FSBs in a VM for VC\\
\hline
$N_o$ & the average number  of FSBs  in a VM for VO\\
\hline
A(r) & the availability of an FSB  with replication degree $r$\\
\hline
\end{tabular}
\caption{Modeling  parameters}
\label{tab:symbol}
\end{table}

\subsection{Impact on Deduplication Efficiency}
Choosing the value $k$ for the most popular chunks affects the deduplication efficiency.
We analyze this impact based on the characteristics  of the VM snapshot traces
studied from  application datasets.
A previous study shows that the popularity of data chunks after local deduplication follows 
a Zipf-like distribution\cite{Breslau1999a} and its
exponent $\alpha$ is ranged between 0.65  and  0.7~\cite{WeiZhangIEEE}. 
Figure~\ref{fig:Datazipf} illustrates the Zipf-like distribution of chunk popularity.
The parameters we will use in our analysis below are defined in
Table~\ref{tab:symbol}.

@@

In Figure~\ref{fig:share} we build a bipartite graph representing the
association from unique file system blocks to their corresponding VMs in the VC
and VO approaches. We use an FSB rather than data chunk as our unit of failure
because the DFS keeps file system blocks as its base unit of storage, and in
the case of 64MB FSBs and 4KB chunks, there are 16K chunks per block. In VC
we assign extra replication to the PDS blocks, while in VO the replication
degree is fixed.

We could consider a third approach (call it VO+), adopting the VO approach, and
then relying on filesystem features to provide extra replication to popular
blocks, without the extra complexity of separating popular and non-popular
data. The problem with this VO+ approach is that without separating the
popular chunks from the less-popular, the popular chunks are dispersed across
all of the filesystem blocks in the storage system. Because the popular blocks
are dispersed, there are very few popular FSBs. The end result is that we
either have to add extra replication on unpopular chunks to ensure the popular
chunks have sufficient replication, or we have to accept a reduced replication
factor on popular chunks. This explains why we must separate the PDS and
non-PDS data in our system to acheive the desired $r_c/r$ ratio (though both can be stored on the same DFS).

@@

Now we  assess  the impact of losing $d$ machines 
to the VC and VO approaches.  
A large $r_c/r$ ratio can have a positive impact on full availability of VM snapshot blocks.
To compute the full availability of all snapshots of a VM, we derive
the probability of losing a snapshot FSB of a VM by
estimating the number of file system blocks per VM in each approach.
An association edge is  drawn  from an FSB to a VM 
if this block is used by the VM.

@@

\item
When $r_c \leq d$, both non-PDS and PDS file system blocks in VC can have a loss.
The full snapshot availability of  a VM in the VC approach is
%\begin{equation}
%\label{eq:VC2}
% (1-\frac{ \binom{d}{r}} { \binom{p}{r} })^{N_1} 
% *
% (1-\frac{ \binom{d}{r_c}} { \binom{p}{r_c} })^{N_2}.
\[
A(r)^{N_1} * A(r_c)^{N_2}.
\]
%\end{equation}
%comparing  Formula~\ref{eq:VO} and~\ref{eq:VC}. 
\end{itemize} 
%In the evaluation discussed in Section~\ref{sect:evaldedup},
We have considered a worst case scenario that
every PDS FSB is shared by all VMs in the VC approach, which leads to a large $N_2$ value. 
Even with that, the availability of VC snapshots is still much higher than VO and  
there are two reasons for this:  1) $N_1$ is much smaller than $N_o$ as discussed previously.
2)  $A(r) < A(r_c)$ because $r < r_c$.  
Table~\ref{tab:fsb-availability} lists the $A(r)$ values with
%that the availability of an individual file system block
different replication degrees, to demonstrate the gap between  $A(r)$ and  $A(r_c)$.
